A Bayesian Perspective \\\large{Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest}

### 2023-02-13 manuscript/manuscript.qmd
file for Circulation

---
title: A Bayesian Perspective 
subtitle: Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest
author:
  - name: James M Brophy MD PhD
    email: james.brophy@mcgill.ca
    affiliations: 
        - id: McGill University 
          name: McGill University Health Center
          department: Centre for Health Outcomes Research (CORE) 
          address: 5252 Boul. de Maisonneuve West Room 2B.37
          city: Montreal
          state: QC
          postal-code: H4A 3S5
    attributes:
        corresponding: true
    note: JMB is a research scholar supported by Les Fonds de Recherche Québec Santé
abstract: |
  A recent randomized clinical trial, INCEPTION, reported in patients with refractory out-of-hospital cardiac arrest, extracorporeal CPR (eCPR) and conventional CPR (cCPR) had similar effects on survival with a favorable neurologic outcome. The current study examines if a Bayesian perspective provides  additional quantative insights. Depending on the prior selected, the Bayesian approach for the INCEPTION intention-to treat (ITT) analysis shows an equivalence probability between 13.4 - 16.8% (defined as 1 / 1.1 < odds ratio (OR) < 1.1). The probability of clinical superiority with eCPR ranges from 65.7 - 77.0 % (defined as OR > 1.1). A similar analyses using INCEPTION per protocol (PP) data shows an equivalence probability between 4.7 - 20.2% with reduced probabilities of clinical superiority not exceeding 25%. It is concluded that a Bayesian perspective allows considerable additional quantative insights into the trial analysis and interpretation. The non-negligible probabilities of increased survival, or even harm, with their considerable residual uncertainties, suggests that additional studies are required before concluding that eCPR and cCPR have similar average survival effects.
keywords: 
  - extracorporeal CPR
  - Bayesian statistics
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: To be determined
      formatting: preprint
      model: 3p
      cite-style: super
editor: 
  markdown: 
    wrap: 72
---

@@

Therefore, in addition to the current data summarized by the probability
of the data (likelihood function), prior probability
distributions are required. Because our main focus is the analysis and 
interpretation of the INCEPTION trial alone, our primary analysis used a default vague parameter prior
($log(\theta) \sim Normal [0, 2.50]$, thereby assuring that the
posterior distribution is dominated by the observed INCEPTION data.

The robustness of the Bayesian approach is often assessed by sensitivity
analyses that examine the variation in the posterior probability as a
function of the choice of different prior distributions. Incorporating prior
information underscores another important advantage of Bayesian
analyses, the ability to learn sequentially. There were two previous RCTs examining
extracorporeal CPR[@RN6759; @RN6751] and while the protocols are not
identical, it may be reasonable to allow this data to serve as 
informed priors for the eCPR parameter, which can be updated with the INCEPTION data.

@@

ITT data from the INCEPTION trial[@CPR2023a] and two other pertinent trials[@RN6759;@RN6751] that also randomized out of hospital cardiac arrest patients to cCPR to eCPR are shown in Table 1. Performing a Bayesian analysis on the INCEPTION trial, using a default vague prior, produces an odds ratio (OR) 1.32 with 95% CrI 0.54 - 3.22. The closeness of this result to the original analysis (OR, 1.4; 95% CI 0.5 - 3.5) confirms the minimal impact of the default vague prior and reveals a Bayesian analysis completely dominated by the observed INCEPTION data.

One of the advantages of a Bayesian approach is the ability to make direct probability statements about the estimand of improved eCPR survival. The eCPR probability density function for improved survival from INCEPTION data with the default vague prior is displayed in Figure 1 and reveals that the probability of enhanced survival with eCPR is `r round(fig_prob*100,1)`%. The probability that the improved survival exceeds a 10% improvement is `r round(fig_prob_1.1*100,1)`% and the ROPE probability is `r round(vague_eq*100,1)`% (Table 2).

@@

Three different informative priors were considered i)  a N(`r round(comb_m_eCPR,2)`, `r round(comb_sd_eCPR,2)`) ii) a N(`r round(enth_m_eCPR,2)`, `r round(enth_sd_eCPR,2)`) iii) a N(`r round(skep_m_eCPR,2)`, `r round(skep_sd_eCPR,2)`) distributions to represent all the combined available RCT data[@RN6759;@RN6751], only the ARREST[@RN6751] data  and only the PRAGUE[@RN6759] data, respectively. These different prior probabilities were updated using with the INCEPTION ITT data to create the posterior distributions displayed in Table 2. The posteriorprobability for enhanced eCPR survival has increased to 80.4% with a skeptical prior, to 84.9% with the enthusiastic prior and, as expected, the associated uncertainty has been reduced, as reflected by the narrower 95% CrI, with the additional data.

@@

INCEPTION did not report a per-protocol analysis. Such an analysis may be helpful in assessing treatment efficacy, unfortunately at the risk of an increased risk of bias by not respecting the ITT principle. From INCEPTION Figure S4[@CPR2023a], it appears that the per protocol data for mortality is 13 survivors from 61 patients in the cCPR group compared to 5 survivors from 46 patients receiving eCPR. With a vague prior, the OR of increased survival with eCPR compared to cCPR is decreased but with very wide CrI (OR 0.45, 95% CrI 0.15 - 1.35), limiting any definitive conclusions. The decreased eCPR success rates in the INCEPTION PP data results in reduced posterior probabilities of eCPR benefit and an increased probability of equivalence or benefit with cCPR. The probabilities of eCPR survival compared to cCPR using the per protocol data and incorporating the previously identified informative priors are shown in Table 3.

@@

This Bayesian analysis of the INCEPTION ITT data alone (analysis with a vague prior) suggests the presence of a non-negligible mortality probability benefit with eCPR, in the range of 73%, compared to cCPR. The probability that this improvement in survival exceeds 10% is almost 66%. The consideration of pertinent prior information from two previous comparable RCTs[@RN6751;@RN6759] results in estimates of improved eCPR survival probabilities, with intact neurological status, in the vicinity of 80-85%. Moreover depending on the choice of prior beliefs included, the posterior probability of at least a 10% improved eCPR survival is in the range of 71 - 77% probability. In contrast to the INCEPTION original conclusion of similar survival effects between the two treatments, this re-analysis suggests only a modest probability of approximately 15% that survival probabilities between the two techniques are within a reasonably acceptable equivalency range. Additional uncertainty about the benefits or harms of eCPR do arise with the per protocol analysis, both with and without informative priors.

The INCEPTION researchers have addressed an important clinical question in the most challenging of research environments and are to be congratulated on their trial design, its execution, and a nuanced discussion. However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the quantitative appreciation of their data, and prevents a full and comprehensive exploitation and updating of past knowledge. For trials such as INCEPTION that fail to meet statistical significance and are often incorrectly thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765]. Rather than dichotomizing results into statistical significance or not, with an obligatory loss of information and understanding, Bayesian analysis concentrates on estimation of key outcome differences with direct probability measures of their uncertainty.

Although this  Bayesian analysis presents the strengths noted above, it also had limitations. First, we did not have access to 
individual data from any of the trials limiting the possibility of examining specific subgroups. Also for the per protocol analysis, this data was only available for survival and not the primary INCEPTION outcome that included survival with intact neurological status. The threshold choices for superiority and equivalence may be seen as arbitrary, although the analyses can be easily repeated for different choices. Given the existence of only 3 trials  estimating the very divergent baseline cCPR rates could not be reliably modeled. We choose not to perform a traditional Bayesian meta-analysis with
a semi-informative prior for the between-study variation  because our
goal was to illustrate how Bayesian principles could be informative
when applied to the analysis and interpretation of new trial data, in situations both with and without previous knowledge.

>>>>

---
title: A Bayesian Perspective 
subtitle: Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest
author:
  - name: James M Brophy MD PhD
    email: james.brophy@mcgill.ca
    affiliations: 
        - id: McGill University 
          name: McGill University Health Center
          department: Centre for Health Outcomes Research (CORE) 
          address: 5252 Boul. de Maisonneuve West Room 2B.37
          city: Montreal
          state: QC
          postal-code: H4A 3S5
    attributes:
        corresponding: true
    note: JMB is a research scholar supported by Les Fonds de Recherche Québec Santé
abstract: |
  A recent randomized clinical trial, INCEPTION, reported in patients with refractory out-of-hospital cardiac arrest, extracorporeal CPR (eCPR) and conventional CPR (cCPR) had similar effects on survival with a favorable neurologic outcome. The current study examines if a Bayesian perspective provides  additional quantative insights. Depending on the prior selected, the Bayesian approach for the INCEPTION intention-to treat (ITT) analysis shows an equivalence probability between 13.4 - 16.8% (defined as 1 / 1.1 < odds ratio (OR) < 1.1). The probability of clinical superiority with eCPR ranges from 65.7 - 77.0 % (defined as OR > 1.1). A similar analyses using INCEPTION per protocol (PP) data shows an equivalence probability between 4.7 - 20.2% with reduced probabilities of clinical superiority not exceeding 25%. It is concluded that a Bayesian perspective allows considerable additional quantative insights into the trial analysis and interpretation. The non-negligible probabilities of increased survival, or even harm, with their considerable residual uncertainties, suggests that additional studies are required before concluding that eCPR and cCPR have similar average survival effects.
keywords: 
  - extracorporeal CPR
  - Bayesian statistics
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Circulation
      formatting: preprint
      model: 3p
      cite-style: super
editor: 
  markdown: 
    wrap: 72
---

@@

Therefore, in addition to the current data summarized by the probability
of the data (likelihood function), prior probability
distributions are required. Because our main focus is the analysis and 
interpretation of the INCEPTION trial[@CPR2023a] alone, our primary analysis used a default vague parameter prior
($log(\theta) \sim Normal [0, 2.50]$, thereby assuring that the
posterior distribution is dominated by the observed INCEPTION[@CPR2023a] data.

The robustness of the Bayesian approach is often assessed by sensitivity
analyses that examine the variation in the posterior probability as a
function of the choice of different prior distributions. Incorporating prior
information underscores another important advantage of Bayesian
analyses, the ability to learn sequentially. There were two previous RCTs examining
extracorporeal CPR[@RN6759; @RN6751] and while the protocols are not
identical, it may be reasonable to allow this data to serve as 
informed priors for the eCPR parameter, which can be updated with the INCEPTION[@CPR2023a] data.

@@

ITT data from the INCEPTION trial[@CPR2023a] and two other pertinent trials[@RN6759;@RN6751] that also randomized out of hospital cardiac arrest patients to cCPR to eCPR are shown in Table 1. Performing a Bayesian analysis on the INCEPTION[@CPR2023a] trial, using a default vague prior, produces an odds ratio (OR) 1.32 with 95% CrI 0.54 - 3.22. The closeness of this result to the original analysis (OR, 1.4; 95% CI 0.5 - 3.5) confirms the minimal impact of the default vague prior and reveals a Bayesian analysis completely dominated by the observed INCEPTION[@CPR2023a] data.

One of the advantages of a Bayesian approach is the ability to make direct probability statements about the estimand of improved eCPR survival. The eCPR probability density function for improved survival from INCEPTION[@CPR2023a] data with the default vague prior is displayed in Figure 1 and reveals that the probability of enhanced survival with eCPR is `r round(fig_prob*100,1)`%. The probability that the improved survival exceeds a 10% improvement is `r round(fig_prob_1.1*100,1)`% and the ROPE probability is `r round(vague_eq*100,1)`% (Table 2).

@@

Three different informative priors were considered i)  a N(`r round(comb_m_eCPR,2)`, `r round(comb_sd_eCPR,2)`) ii) a N(`r round(enth_m_eCPR,2)`, `r round(enth_sd_eCPR,2)`) iii) a N(`r round(skep_m_eCPR,2)`, `r round(skep_sd_eCPR,2)`) distributions to represent all the combined available RCT data[@RN6759;@RN6751], only the ARREST[@RN6751] data  and only the PRAGUE[@RN6759] data, respectively. These different prior probabilities were updated using with the INCEPTION[@CPR2023a] ITT data to create the posterior distributions displayed in Table 2. The posteriorprobability for enhanced eCPR survival has increased to 80.4% with a skeptical prior, to 84.9% with the enthusiastic prior and, as expected, the associated uncertainty has been reduced, as reflected by the narrower 95% CrI, with the additional data.

@@

INCEPTION[@CPR2023a] did not report a per-protocol analysis. Such an analysis may be helpful in assessing treatment efficacy, unfortunately at the risk of an increased risk of bias by not respecting the ITT principle. From INCEPTION[@CPR2023a] Figure S4[@CPR2023a], it appears that the per protocol data for mortality is 13 survivors from 61 patients in the cCPR group compared to 5 survivors from 46 patients receiving eCPR. With a vague prior, the OR of increased survival with eCPR compared to cCPR is decreased but with very wide CrI (OR 0.45, 95% CrI 0.15 - 1.35), limiting any definitive conclusions. The decreased eCPR success rates in the INCEPTION[@CPR2023a] per protocol data results in reduced posterior probabilities of eCPR benefit and an increased probability of equivalence or benefit with cCPR. The probabilities of eCPR survival compared to cCPR using the per protocol data and incorporating the previously identified informative priors are shown in Table 3.

@@

This Bayesian analysis of the INCEPTION[@CPR2023a] ITT data alone (analysis with a vague prior) suggests the presence of a non-negligible mortality probability benefit with eCPR, in the range of 73%, compared to cCPR. The probability that this improvement in survival exceeds 10% is almost 66%. The consideration of pertinent prior information from two previous comparable RCTs[@RN6751;@RN6759] results in estimates of improved eCPR survival probabilities, with intact neurological status, in the vicinity of 80-85%. Moreover depending on the choice of prior beliefs included, the posterior probability of at least a 10% improved eCPR survival is in the range of 71 - 77% probability. In contrast to the INCEPTION[@CPR2023a] original conclusion of similar survival effects between the two treatments, this re-analysis suggests only a modest probability of approximately 15% that survival probabilities between the two techniques are within a reasonably acceptable equivalency range. Additional uncertainty about the benefits or harms of eCPR do arise with the per protocol analysis, both with and without informative priors.

The INCEPTION[@CPR2023a] researchers have addressed an important clinical question in the most challenging of research environments and are to be congratulated on their trial design, its execution, and a nuanced discussion. However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the quantitative appreciation of their data, and prevents a full and comprehensive exploitation and updating of past knowledge. For trials such as INCEPTION[@CPR2023a] that fail to meet statistical significance and are often incorrectly thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765]. Rather than dichotomizing results into statistical significance or not, with an obligatory loss of information and understanding, Bayesian analysis concentrates on estimation of key outcome differences with direct probability measures of their uncertainty.

Although this  Bayesian analysis presents the strengths noted above, it also had limitations. First, we did not have access to 
individual data from any of the trials limiting the possibility of examining specific subgroups. Also for the per protocol analysis, this data was only available for survival and not the primary INCEPTION[@CPR2023a] outcome that included survival with intact neurological status. The threshold choices for superiority and equivalence may be seen as arbitrary, although the analyses can be easily repeated for different choices. Given the existence of only 3 trials  estimating the very divergent baseline cCPR rates could not be reliably modeled. We choose not to perform a traditional Bayesian meta-analysis with
a semi-informative prior for the between-study variation  because our
goal was to illustrate how Bayesian principles could be informative
when applied to the analysis and interpretation of new trial data, in situations both with and without previous knowledge.

### 2023-02-13 manuscript/manuscript.tex
file for Circulation

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{To be determined}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A Bayesian Perspective},
  pdfauthor={James M Brophy MD PhD},
  pdfkeywords={extracorporeal CPR, Bayesian statistics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, sharp corners, enhanced, frame hidden, breakable]}{\end{tcolorbox}}\fi

@@

Therefore, in addition to the current data summarized by the probability
of the data (likelihood function), prior probability distributions are
required. Because our main focus is the analysis and interpretation of
the INCEPTION trial alone, our primary analysis used a default vague
parameter prior (\(log(\theta) \sim Normal [0, 2.50]\), thereby assuring
that the posterior distribution is dominated by the observed INCEPTION
data.

The robustness of the Bayesian approach is often assessed by sensitivity
analyses that examine the variation in the posterior probability as a
function of the choice of different prior distributions. Incorporating
prior information underscores another important advantage of Bayesian
analyses, the ability to learn sequentially. There were two previous
RCTs examining extracorporeal CPR\citep{RN6759, RN6751} and while the
protocols are not identical, it may be reasonable to allow this data to
serve as informed priors for the eCPR parameter, which can be updated
with the INCEPTION data.

@@

ITT data from the INCEPTION trial\citep{CPR2023a} and two other
pertinent trials\citep{RN6759, RN6751} that also randomized out of
hospital cardiac arrest patients to cCPR to eCPR are shown in Table 1.
Performing a Bayesian analysis on the INCEPTION trial, using a default
vague prior, produces an odds ratio (OR) 1.32 with 95\% CrI 0.54 - 3.22.
The closeness of this result to the original analysis (OR, 1.4; 95\% CI
0.5 - 3.5) confirms the minimal impact of the default vague prior and
reveals a Bayesian analysis completely dominated by the observed
INCEPTION data.

One of the advantages of a Bayesian approach is the ability to make
direct probability statements about the estimand of improved eCPR
survival. The eCPR probability density function for improved survival
from INCEPTION data with the default vague prior is displayed in Figure
1 and reveals that the probability of enhanced survival with eCPR is
72.7\%. The probability that the improved survival exceeds a 10\%
improvement is 65.7\% and the ROPE probability is 13.4\% (Table 2).

Three different informative priors were considered i) a N(0.32, 0.47)
ii) a N(0.43, 0.49) iii) a N(0.31, 0.46) distributions to represent all
the combined available RCT data\citep{RN6759, RN6751}, only the
ARREST\citep{RN6751} data and only the PRAGUE\citep{RN6759} data,
respectively. These different prior probabilities were updated using
with the INCEPTION ITT data to create the posterior distributions
displayed in Table 2. The posteriorprobability for enhanced eCPR
survival has increased to 80.4\% with a skeptical prior, to 84.9\% with
the enthusiastic prior and, as expected, the associated uncertainty has
been reduced, as reflected by the narrower 95\% CrI, with the additional
data.

@@

INCEPTION did not report a per-protocol analysis. Such an analysis may
be helpful in assessing treatment efficacy, unfortunately at the risk of
an increased risk of bias by not respecting the ITT principle. From
INCEPTION Figure S4\citep{CPR2023a}, it appears that the per protocol
data for mortality is 13 survivors from 61 patients in the cCPR group
compared to 5 survivors from 46 patients receiving eCPR. With a vague
prior, the OR of increased survival with eCPR compared to cCPR is
decreased but with very wide CrI (OR 0.45, 95\% CrI 0.15 - 1.35),
limiting any definitive conclusions. The decreased eCPR success rates in
the INCEPTION PP data results in reduced posterior probabilities of eCPR
benefit and an increased probability of equivalence or benefit with
cCPR. The probabilities of eCPR survival compared to cCPR using the per
protocol data and incorporating the previously identified informative
priors are shown in Table 3.

@@

This Bayesian analysis of the INCEPTION ITT data alone (analysis with a
vague prior) suggests the presence of a non-negligible mortality
probability benefit with eCPR, in the range of 73\%, compared to cCPR.
The probability that this improvement in survival exceeds 10\% is almost
66\%. The consideration of pertinent prior information from two previous
comparable RCTs\citep{RN6751, RN6759} results in estimates of improved
eCPR survival probabilities, with intact neurological status, in the
vicinity of 80-85\%. Moreover depending on the choice of prior beliefs
included, the posterior probability of at least a 10\% improved eCPR
survival is in the range of 71 - 77\% probability. In contrast to the
INCEPTION original conclusion of similar survival effects between the
two treatments, this re-analysis suggests only a modest probability of
approximately 15\% that survival probabilities between the two
techniques are within a reasonably acceptable equivalency range.
Additional uncertainty about the benefits or harms of eCPR do arise with
the per protocol analysis, both with and without informative priors.

The INCEPTION researchers have addressed an important clinical question
in the most challenging of research environments and are to be
congratulated on their trial design, its execution, and a nuanced
discussion. However the natural constraints of standard statistical
analyses, also known as null hypothesis significance testing, limits the
quantitative appreciation of their data, and prevents a full and
comprehensive exploitation and updating of past knowledge. For trials
such as INCEPTION that fail to meet statistical significance and are
often incorrectly thought of as ``negative'' trials, null hypothesis
significance testing favors confusion ``between absence of evidence and
evidence of absence''\citep{RN6765}. Rather than dichotomizing results
into statistical significance or not, with an obligatory loss of
information and understanding, Bayesian analysis concentrates on
estimation of key outcome differences with direct probability measures
of their uncertainty.

Although this Bayesian analysis presents the strengths noted above, it
also had limitations. First, we did not have access to individual data
from any of the trials limiting the possibility of examining specific
subgroups. Also for the per protocol analysis, this data was only
available for survival and not the primary INCEPTION outcome that
included survival with intact neurological status. The threshold choices
for superiority and equivalence may be seen as arbitrary, although the
analyses can be easily repeated for different choices. Given the
existence of only 3 trials estimating the very divergent baseline cCPR
rates could not be reliably modeled. We choose not to perform a
traditional Bayesian meta-analysis with a semi-informative prior for the
between-study variation because our goal was to illustrate how Bayesian
principles could be informative when applied to the analysis and
interpretation of new trial data, in situations both with and without
previous knowledge.

>>>>

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Circulation}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A Bayesian Perspective},
  pdfauthor={James M Brophy MD PhD},
  pdfkeywords={extracorporeal CPR, Bayesian statistics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

Therefore, in addition to the current data summarized by the probability
of the data (likelihood function), prior probability distributions are
required. Because our main focus is the analysis and interpretation of
the INCEPTION trial\citep{CPR2023a} alone, our primary analysis used a
default vague parameter prior (\(log(\theta) \sim Normal [0, 2.50]\),
thereby assuring that the posterior distribution is dominated by the
observed INCEPTION\citep{CPR2023a} data.

The robustness of the Bayesian approach is often assessed by sensitivity
analyses that examine the variation in the posterior probability as a
function of the choice of different prior distributions. Incorporating
prior information underscores another important advantage of Bayesian
analyses, the ability to learn sequentially. There were two previous
RCTs examining extracorporeal CPR\citep{RN6759, RN6751} and while the
protocols are not identical, it may be reasonable to allow this data to
serve as informed priors for the eCPR parameter, which can be updated
with the INCEPTION\citep{CPR2023a} data.

@@

ITT data from the INCEPTION trial\citep{CPR2023a} and two other
pertinent trials\citep{RN6759, RN6751} that also randomized out of
hospital cardiac arrest patients to cCPR to eCPR are shown in Table 1.
Performing a Bayesian analysis on the INCEPTION\citep{CPR2023a} trial,
using a default vague prior, produces an odds ratio (OR) 1.32 with 95\%
CrI 0.54 - 3.22. The closeness of this result to the original analysis
(OR, 1.4; 95\% CI 0.5 - 3.5) confirms the minimal impact of the default
vague prior and reveals a Bayesian analysis completely dominated by the
observed INCEPTION\citep{CPR2023a} data.

One of the advantages of a Bayesian approach is the ability to make
direct probability statements about the estimand of improved eCPR
survival. The eCPR probability density function for improved survival
from INCEPTION\citep{CPR2023a} data with the default vague prior is
displayed in Figure 1 and reveals that the probability of enhanced
survival with eCPR is 72.7\%. The probability that the improved survival
exceeds a 10\% improvement is 65.7\% and the ROPE probability is 13.4\%
(Table 2).

Three different informative priors were considered i) a N(0.32, 0.47)
ii) a N(0.43, 0.49) iii) a N(0.31, 0.46) distributions to represent all
the combined available RCT data\citep{RN6759, RN6751}, only the
ARREST\citep{RN6751} data and only the PRAGUE\citep{RN6759} data,
respectively. These different prior probabilities were updated using
with the INCEPTION\citep{CPR2023a} ITT data to create the posterior
distributions displayed in Table 2. The posteriorprobability for
enhanced eCPR survival has increased to 80.4\% with a skeptical prior,
to 84.9\% with the enthusiastic prior and, as expected, the associated
uncertainty has been reduced, as reflected by the narrower 95\% CrI,
with the additional data.

@@

INCEPTION\citep{CPR2023a} did not report a per-protocol analysis. Such
an analysis may be helpful in assessing treatment efficacy,
unfortunately at the risk of an increased risk of bias by not respecting
the ITT principle. From INCEPTION\citep{CPR2023a} Figure
S4\citep{CPR2023a}, it appears that the per protocol data for mortality
is 13 survivors from 61 patients in the cCPR group compared to 5
survivors from 46 patients receiving eCPR. With a vague prior, the OR of
increased survival with eCPR compared to cCPR is decreased but with very
wide CrI (OR 0.45, 95\% CrI 0.15 - 1.35), limiting any definitive
conclusions. The decreased eCPR success rates in the
INCEPTION\citep{CPR2023a} per protocol data results in reduced posterior
probabilities of eCPR benefit and an increased probability of
equivalence or benefit with cCPR. The probabilities of eCPR survival
compared to cCPR using the per protocol data and incorporating the
previously identified informative priors are shown in Table 3.

@@

This Bayesian analysis of the INCEPTION\citep{CPR2023a} ITT data alone
(analysis with a vague prior) suggests the presence of a non-negligible
mortality probability benefit with eCPR, in the range of 73\%, compared
to cCPR. The probability that this improvement in survival exceeds 10\%
is almost 66\%. The consideration of pertinent prior information from
two previous comparable RCTs\citep{RN6751, RN6759} results in estimates
of improved eCPR survival probabilities, with intact neurological
status, in the vicinity of 80-85\%. Moreover depending on the choice of
prior beliefs included, the posterior probability of at least a 10\%
improved eCPR survival is in the range of 71 - 77\% probability. In
contrast to the INCEPTION\citep{CPR2023a} original conclusion of similar
survival effects between the two treatments, this re-analysis suggests
only a modest probability of approximately 15\% that survival
probabilities between the two techniques are within a reasonably
acceptable equivalency range. Additional uncertainty about the benefits
or harms of eCPR do arise with the per protocol analysis, both with and
without informative priors.

The INCEPTION\citep{CPR2023a} researchers have addressed an important
clinical question in the most challenging of research environments and
are to be congratulated on their trial design, its execution, and a
nuanced discussion. However the natural constraints of standard
statistical analyses, also known as null hypothesis significance
testing, limits the quantitative appreciation of their data, and
prevents a full and comprehensive exploitation and updating of past
knowledge. For trials such as INCEPTION\citep{CPR2023a} that fail to
meet statistical significance and are often incorrectly thought of as
``negative'' trials, null hypothesis significance testing favors
confusion ``between absence of evidence and evidence of
absence''\citep{RN6765}. Rather than dichotomizing results into
statistical significance or not, with an obligatory loss of information
and understanding, Bayesian analysis concentrates on estimation of key
outcome differences with direct probability measures of their
uncertainty.

Although this Bayesian analysis presents the strengths noted above, it
also had limitations. First, we did not have access to individual data
from any of the trials limiting the possibility of examining specific
subgroups. Also for the per protocol analysis, this data was only
available for survival and not the primary INCEPTION\citep{CPR2023a}
outcome that included survival with intact neurological status. The
threshold choices for superiority and equivalence may be seen as
arbitrary, although the analyses can be easily repeated for different
choices. Given the existence of only 3 trials estimating the very
divergent baseline cCPR rates could not be reliably modeled. We choose
not to perform a traditional Bayesian meta-analysis with a
semi-informative prior for the between-study variation because our goal
was to illustrate how Bayesian principles could be informative when
applied to the analysis and interpretation of new trial data, in
situations both with and without previous knowledge.
